Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-07-12 prop_letter.tex
prop_letter.tex Updating according to the meeting...

We will proceed by assuming a QIVMP $\bar{\mu}$
and derive a contradiction using the following nine ``Mermin-Peres square''
observables $\mathbf{O}_{ij}$\nb\cite{Mermin1990Simple,peres1995quantum}:
\begin{center}
\begin{tabular}{|c|c|c|}
\hline 
$\mathbb{1}\otimes\sigma_{z}$  & $\sigma_{z}\otimes\mathbb{1}$  & $\sigma_{z}\otimes\sigma_{z}$ \tabularnewline
\hline 
$\sigma_{x}\otimes\mathbb{1}$  & $\mathbb{1}\otimes\sigma_{x}$  & $\sigma_{x}\otimes\sigma_{x}$ \tabularnewline
\hline 
$\sigma_{x}\otimes\sigma_{z}$  & $\sigma_{z}\otimes\sigma_{x}$  & $\sigma_{y}\otimes\sigma_{y}$ \tabularnewline
\hline 
\end{tabular}\,,
\par\end{center}

\noindent The argument refers to the expectation value of these
observables. In the case of QIVPM this is defined as follows.

\begin{definition}Given a Hilbert space of $d\ge3$, for every
  QIVPM\nb$\bar{\mu}:\events\rightarrow\mathscr{I}$, any
  observable\nb$\mathbf{O}$ with spectral decomposition
  $\mathbf{O}=\sum_{i=1}^{N}\lambda_{i}P_{i}$, and any minimal
  sub-event space $\events'$ containing
  $\left\{ P_{i}\right\} _{i=1}^{N}$, we have
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\left[\min_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho},\max_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho}\right]\,.\label{eq:quantum-interval-expectation-core}
\end{equation}
\end{definition}

The expectation value is intuitively between two possible outcomes
which lie between the minimum and maximum bounds of the given
probability intervals. (This definition is consistent with the
classical Choquet integral on commuting sub-event spaces by adopting
the classical
results\nb\cite{GilboaSchmeidler1994,Grabisch2016}.)
If the QIVPM is cryptodeterministic, this interval collapses to a
singleton $\left\{ \lambda\right\} $, where $\lambda$ is an eigenvalue
of~$\mathbf{O}$. If we define the multiplication of two singleton
sets\nb$\left\{ \lambda_{1}\right\} $ and
$\left\{ \lambda_{2}\right\} $ to be
$\left\{ \lambda_{1}\lambda_{2}\right\} $, then we have
\begin{equation}
\prod_{j=1}^{N}\expval{\mathbf{O}_{j}}_{\bar{\mu}}=\expval{\prod_{j=1}^{N}\mathbf{O}_{j}}_{\bar{\mu}}\label{eq:Kochen-Specker-Rules}
\end{equation}
for any sequence of commuting observables\nb$\left\{ \mathbf{O}_{j}\right\} _{j=1}^{N}$.

Since the expectation value\nb$\expval{\cdot}_{\bar{\mu}}$ satisfies
these two properties, the observables in the Mermin-Peres square must
have the the expectation value\nb$\left\{ 1\right\} $ or
$\left\{ -1\right\} $.  In each row and column, the operators commute,
and each operator is the product of the two others, except in the
third column
$\left(\sigma_{z}\otimes\sigma_{z}\right)\left(\sigma_{x}\otimes\sigma_{x}\right)=-\sigma_{y}\otimes\sigma_{y}$.
Therefore, it is clearly impossible to find a QIVPM\nb$\bar{\mu}$ such
that $\langle \mathbf{O}_{ij}\rangle_{\bar \mu}$
%\begin{equation}
%\begin{aligned}\langle\mathbb{1} & \otimes\sigma_{z}\rangle_{\bar{\mu}} & \langle\sigma_{z} & \otimes\mathbb{1}\rangle_{\bar{\mu}} & \langle\sigma_{z} %& \otimes\sigma_{z}\rangle_{\bar{\mu}}\\
%\langle\sigma_{x} & \otimes\mathbb{1}\rangle_{\bar{\mu}} & \langle\mathbb{1} & \otimes\sigma_{x}\rangle_{\bar{\mu}} & \langle\sigma_{x} & %\otimes\sigma_{x}\rangle_{\bar{\mu}}\\
%\langle\sigma_{x} & \otimes\sigma_{z}\rangle_{\bar{\mu}} & \langle\sigma_{z} & \otimes\sigma_{x}\rangle_{\bar{\mu}} & \langle\sigma_{y} & %\otimes\sigma_{y}\rangle_{\bar{\mu}}
%\end{aligned}
%\label{eq:MerminSquare-values}
%\end{equation}
satisfies Eq.~(\ref{eq:Kochen-Specker-Rules}).

The question now is what happens if we relax the perfect measurement
assumption that entails that the intervals are precisely $[0,0]$ and
$[1,1]$. What if the measurement imperfections occasionally cause an
ideally impossible event to be detected or an ideally certain event to
be not be detected. Mermin argues in this case that:
\begin{quote}
  \ldots although the outcomes deduced from such imperfect
  measurements will occasionally differ dramatically from those
  allowed in the ideal case, if the misalignment is very slight, the
  statistical distribution of outcomes will differ only slightly from
  the ideal case~\citep{Mermin1999}.
\end{quote}

>>>>

We will proceed by assuming a cryptodeterministic QIVMP $\bar{\mu}$
and derive a contradiction using the following nine ``Mermin-Peres square''
observables $\mathbf{O}_{ij}$\nb\cite{Mermin1990Simple,peres1995quantum}:
\begin{center}
\begin{tabular}{|c|c|c|}
\hline 
$\mathbb{1}\otimes\sigma_{z}$  & $\sigma_{z}\otimes\mathbb{1}$  & $\sigma_{z}\otimes\sigma_{z}$ \tabularnewline
\hline 
$\sigma_{x}\otimes\mathbb{1}$  & $\mathbb{1}\otimes\sigma_{x}$  & $\sigma_{x}\otimes\sigma_{x}$ \tabularnewline
\hline 
$\sigma_{x}\otimes\sigma_{z}$  & $\sigma_{z}\otimes\sigma_{x}$  & $\sigma_{y}\otimes\sigma_{y}$ \tabularnewline
\hline 
\end{tabular}\,,
\par\end{center}

\noindent where $\left\{\mathbb{1},\sigma_{x},\sigma_{y},\sigma_{z}\right\}$ are
Pauli matrices acting on a pair of spin-$\frac{1}{2}$ state. We now define the
expectation value of the QIVPM for our Mermin-Peres square.

\begin{definition}Given a Hilbert space of dimension\nb$d\ge3$, for every
  QIVPM\nb$\bar{\mu}:\events\rightarrow\mathscr{I}$, any
  observable\nb$\mathbf{O}$ with spectral decomposition
  $\mathbf{O}=\sum_{i=1}^{N}\lambda_{i}P_{i}$, and any minimal
  sub-event space $\events'$ containing
  $\left\{ P_{i}\right\} _{i=1}^{N}$, we have
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\left[\min_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho},\max_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho}\right]\,.\label{eq:quantum-interval-expectation-core}
\end{equation}
\end{definition}

The expectation value is intuitively between two possible outcomes,
which themselves lie between the minimum and maximum bounds of the given
probability intervals. (This definition is consistent with the
classical Choquet integral on commuting sub-event spaces by adopting
the classical
results\nb\cite{GilboaSchmeidler1994,Grabisch2016}.)
If the QIVPM is cryptodeterministic, this interval collapses to a
singleton $\left\{ \lambda\right\} $, where $\lambda$ is an eigenvalue
of~$\mathbf{O}$. If we define the multiplication of two singleton
sets\nb$\left\{ \lambda_{1}\right\} $ and
$\left\{ \lambda_{2}\right\} $ to be
$\left\{ \lambda_{1}\lambda_{2}\right\} $, then we have
\begin{equation}
\prod_{j=1}^{N}\expval{\mathbf{O}_{j}}_{\bar{\mu}}=\expval{\prod_{j=1}^{N}\mathbf{O}_{j}}_{\bar{\mu}}\label{eq:Kochen-Specker-Rules}
\end{equation}
for any set of commuting observables\nb$\left\{ \mathbf{O}_{j}\right\}
_{j=1}^{N}$.

Since the expectation value\nb$\expval{\cdot}_{\bar{\mu}}$ satisfies
these two properties, the observables in the Mermin-Peres square must
have the the expectation value\nb$\left\{ 1\right\} $ or
$\left\{ -1\right\} $.  In each row and column, the operators commute,
and each operator is the product of two others, except in the
third column, where
$\left(\sigma_{z}\otimes\sigma_{z}\right)\left(\sigma_{x}\otimes\sigma_{x}\right)=-\sigma_{y}\otimes\sigma_{y}$.
Therefore, it is clearly impossible to find a QIVPM\nb$\bar{\mu}$ such
that $\langle \mathbf{O}_{ij}\rangle_{\bar \mu}$
%\begin{equation}
%\begin{aligned}\langle\mathbb{1} & \otimes\sigma_{z}\rangle_{\bar{\mu}} & \langle\sigma_{z} & \otimes\mathbb{1}\rangle_{\bar{\mu}} & \langle\sigma_{z} %& \otimes\sigma_{z}\rangle_{\bar{\mu}}\\
%\langle\sigma_{x} & \otimes\mathbb{1}\rangle_{\bar{\mu}} & \langle\mathbb{1} & \otimes\sigma_{x}\rangle_{\bar{\mu}} & \langle\sigma_{x} & %\otimes\sigma_{x}\rangle_{\bar{\mu}}\\
%\langle\sigma_{x} & \otimes\sigma_{z}\rangle_{\bar{\mu}} & \langle\sigma_{z} & \otimes\sigma_{x}\rangle_{\bar{\mu}} & \langle\sigma_{y} & %\otimes\sigma_{y}\rangle_{\bar{\mu}}
%\end{aligned}
%\label{eq:MerminSquare-values}
%\end{equation}
satisfies Eq.~(\ref{eq:Kochen-Specker-Rules}).

The question now is what happens if we relax the perfect measurement
assumption that the intervals are precisely $[0,0]$ and
$[1,1]$. We need to examine what happens if
measurement imperfections occasionally cause an
ideally impossible event to be detected or an ideally certain event to
be not be detected. Mermin argues in this case that:
\begin{quote}
  \ldots although the outcomes deduced from such imperfect
  measurements will occasionally differ dramatically from those
  allowed in the ideal case, if the misalignment is very slight, the
  statistical distribution of outcomes will differ only slightly from
  the ideal case~\citep{Mermin1999}.
\end{quote}
